A Switching Planner for Combined Task and Observation Planning

### 2010-11-26 code/subarchitectures/planner.sa/trunk/src/latex/ICAPS-2011/switching.tex
new intro to swiching section

We describe our continual planning approach that switches between
sequential and contingent sessions. This proceeds by interleaved
planning and execution in a deterministic-action POMDP described in
DTPDDL. During a sequential session, planning is performed by a
``classical'' system,\footnote{That is, a system designed to solve
fully observable deterministic tasks.}  and execution proceeds
according to the {\em trace} computed by that system. Taking the form
of a classical plan, the trace specifies a sequence of POMDP
actions that achieve the agent's objectives in a deterministic
approximation, i.e., {\em determinisation}, of the problem at
hand. More precisely, the trace is an interleaved sequence of POMDP
actions and {\em assumptive} actions. The latter correspond to
assumptions the planner makes about the truth value of {\em runtime
state variables} -- e.g. that a box of cornflakes is located in the
kitchen at the third plan step. Another kind of assumption, called an
{\em applicability} assumption, is made when the trace includes an
action $\action$ that is not applicable with probability $1$ at the
belief-state
\bstate\ that the system is projected to be in when \action\ is
scheduled for execution, i.e., $\exists\state\in\states\;
\bstate(\state) > 0$ and $\poss(\action)\not\subseteq\state$. By
scheduling
\action,  the serial planner makes an assumption about the
observability of the precondition $\poss(\action)$.

Our approach always begins with a sequential session. Non-assumptive
actions from the trace are executed in sequence until the
applicability of the next scheduled action is uncertain. Here, that
action is denoted \switchAction.  A contingent session then begins,
which tailors sensory processing by the agent to determine whether the
assumptions made in the trace hold. For each {\em active} assumption
in the trace, we add a {\em disconfirm} action to the POMDP whose
execution is rewarding (resp. costly) if the assumption is false
(resp. true). We also add one {\em confirm} action that is rewarding
(resp. costly) to execute if $\poss(\action)$ is true
(resp. false). If execution of the contingent plan applies a
disconfirm action, then a new sequential session begins at the {\em
underlying} belief-state. If the confirm action is executed, the
sequential session is resumed, and
\switchAction\ is executed. Otherwise, control rests in the continual
session.
%%
Because contingent planning is only practical in relatively small
POMDPs, contingent sessions proceed in an abstract decision process
determined by the current trace. That abstract process is constructed
by first omitting all variables from the model that are not featured
in the trace, and by then iteratively refining that abstraction while
the resulting model is of a practicable size.

Finally, whether proceeding in a sequential or contingent session, our
continual planner maintains a factored representation of successive
{\em belief-states}, by performing belief revision according to
Eq~\ref{eq:revision}. Moreover, our internal representation of that
distribution corresponds more-or-less directly to an $(\pp{:init})$
declaration in DTPDDL. Our approach uses that underlying belief-state
as: (1) the source of determinisations for sequential planning, (2)
the basis for switching, and (3) a mechanism to guide construction of
an abstract process for contingent sessions.

>>>>

We describe our continual planning approach that switches between
sequential and contingent sessions. It proceeds by interleaved
planning and execution in a deterministic-action POMDP described in
DTPDDL. During a sequential session, planning is performed by a
``classical'' system,\footnote{That is, a planner designed to solve
fully observable deterministic tasks.}  and execution proceeds
according to the {\em trace} computed by that system. Taking the form
of a classical plan, the trace specifies a sequence of POMDP actions
that achieve the agent's objectives in a deterministic approximation,
i.e., {\em determinisation}, of the problem at hand. More precisely,
the trace is an interleaved sequence of POMDP actions and {\em
assumptive} actions. The latter correspond to assumptions the planner
makes about the truth value of propositions -- e.g. that a box of
cornflakes is located in the kitchen at the third plan step. Another
kind of assumption, called an {\em applicability} assumption, is made
when the trace includes an action $\action$ that is not applicable
with probability $1$ at the belief-state
\bstate\ that the system is projected to be in when \action\ is
scheduled for execution, i.e., $\exists\state\in\states\;
\bstate(\state) > 0$ and $\poss(\action)\not\subseteq\state$. By
scheduling
\action,  the serial planner makes an assumption about the
observability of the precondition $\poss(\action)$.

Our approach always begins with a sequential session. Non-assumptive
actions from the trace are executed in sequence until the
applicability of the next scheduled action is uncertain. We denote
that action \switchAction.  A contingent session then begins, which
tailors sensory processing by the agent to determine whether the
assumptions made in the trace hold. For assumptions in the trace, we
add a {\em disconfirm} action to the POMDP whose execution is
rewarding (resp. costly) if the assumption is false (resp. true). We
also add one {\em confirm} action that is rewarding (resp. costly) to
execute if $\poss(\switchAction)$ is true (resp. false). If execution
of the contingent plan applies a disconfirm action, then a new
sequential session begins at the {\em underlying} belief-state. If the
confirm action is executed, the sequential session is resumed, and
\switchAction\ is executed. Otherwise, control rests in the continual
session.
%%
Because contingent planning is only practical in relatively small
POMDPs, contingent sessions plan in an abstract decision process
determined by the current trace. This abstraction is constructed by
first omitting all propositions that do not featured in the trace, and
by then iteratively refining the model while the result is of a
practicable size.

Finally, whether proceeding in a sequential or contingent session, our
continual planner maintains a factored representation of successive
belief-states, by performing belief revision according to
Eq~\ref{eq:revision}. Moreover, our internal representation the
underlying belief-state corresponds more-or-less directly to an
$(\pp{:init})$ declaration in DTPDDL. Our approach uses that
distribution: (1) as the source of candidate determinisations for
sequential planning, (2) to determine when to switch, and (3) as a
mechanism to guide construction of an abstract process for contingent
sessions.
